Lecture 8: Coalition Games

### 2014-11-07 course-game/session8/ehsan_session1.tex
changes

\documentclass{beamer}
\usepackage[latin1]{inputenc,colortbl}
\usepackage{epsfig}
\usetheme{Frankfurt}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[page number]
%\newtheorem{definition}{Definition}
\title[Lecture 8: Extensive Form Games with Perfect Information]{Lecture 8: Extensive Form Games with Perfect Information}
\author{INSE 6441/2 DD\\ \vspace{0.2cm} Applied Game Theory and Mechanism Design \\ Fall 2014}
\institute{Department of Computer Science and Software Engineering\\Concordia University}
\date{Novermber 6, 2014}
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame1 title page %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\titlepage
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame2 outline page %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Overview}
    \begin{itemize}
     	\itemsep=.5cm
    	\item {\bf Coalition Games}
    	\item Stability and Core
    	\item Fairness and Shapley Value
    	\item Other Solution Conecpts
        \item Practical Example
    \end{itemize}
\end{frame}

@@

\begin{frame}{Cooperative Game Theory}
  \begin{itemize}
     \item Cooperative games are a branch of game theory that models cooperation or collaboration between agents.
     \begin{itemize}
        \item Cooperation to perform a set of tasks that requires different expertise.
        \item Agents do not have enough resource on their own to perform the tasks.
        \item Examples:
        \begin{itemize}
            \item Robots have the ability to move objects in a plant, but multiple robots are required to move a heavy box.
            \item Transportation domain: agents are trucks, trains, airplanes, ships... a task is a good to be transported.
        \end{itemize}
    \end{itemize}
    \item \textbf{Issues:}
        \begin{itemize}
            \item Coalition formation.
            \item Rewarding members when a task is completed.
        \end{itemize}
  \end{itemize}
\end{frame}

@@

    $\forall C \subseteq S \subseteq N$
    \begin{itemize}
        \item {\color{blue} Monotonic:} $v(C) \leq v(S)$
    \end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Solution Properties}
    Let $x \in R^n$ be a solution of the coalition game $(N,v)$
    \begin{itemize}
        \item {\color{blue} Feasible solution:} $\sum_{i \in N} x(i) \leq v(N)$
        \item {\color{blue} Efficiency:} $\sum_{i \in N} x(i) = v(N)$
        \begin{itemize}
            \item the payoff distribution is an allocation of the entire worth of the grand coalition to all agents.
        \end{itemize}
        \item {\color{blue} Individual rationality:} $\forall i \in N, x(i) \geq v(\{i\})$
        \begin{itemize}
            \item player obtains at least its self-value of payoff.
        \end{itemize}
        \item {\color{blue} Group rationality:} $\forall C \subseteq N, \sum_{i \in N} x(i) \geq v(C)$
        \begin{itemize}
            \item if $\sum_{i \in N} x(i) < v(C)$ is not worth it for some $C$
            \item if $\sum_{i \in N} x(i) > v(C)$ is not possible for all $C$
        \end{itemize}
        \item {\color{blue} Pareto Optimal:} $\sum_{i \in N} x(i) = v(N)$
        \begin{itemize}
            \item no agent can improve its payoff without lowering the payoff of another agent
        \end{itemize}
        \item {\color{blue} Anonymity:} a solution is independent of the names of the player.
    \end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Overview}
    \begin{itemize}
     	\itemsep=.5cm
    	\item Coalition Games
    	\item {\bf Stability and Core}
    	\item Fairness and Shapley Value
    	\item Other Solution Conecpts
        \item Practical Example
    \end{itemize}
\end{frame}

@@

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} {Example 1: Vector game}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} {Example 2: Three-player majority game}
    \begin{center}
      $N = \{1,2,3\}$ \\
      $v(\{i\}) = 0,$\\
      $v(\{C\}) = 0, for |C| = 2$ \\
      $v(N) = 1$ \\
    \end{center}
    \[ (x_1, x_2, x_3) \in Core(N,n) \iff \left\{
      \begin{array}{l l}
        x_i \geq 0 & \quad \forall i \in N\\
        x_i + x_j \geq \alpha & \quad \forall(i,j) \in N^2 i \neq j\\
        \sum_{i \in N} x_i = 1 &
      \end{array} \right.\]
    \[ (x_1, x_2, x_3) \in Core(N,n) \iff \left\{
      \begin{array}{l l}
        0 \leq x_i \leq 1 - \alpha & \quad \forall i \in N\\
        \sum_{i \in N} x_i = 1 &
      \end{array} \right.\]

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Overview}
    \begin{itemize}
     	\itemsep=.5cm
    	\item Coalition Games
    	\item Stability and Core
    	\item {\bf Fairness and Shapley Value}
    	\item Other Solution Conecpts
        \item Practical Example
    \end{itemize}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Overview}
    \begin{itemize}
     	\itemsep=.5cm
    	\item Coalition Games
    	\item Stability and Core
    	\item Fairness and Shapley Value
    	\item {\bf Other Solution Conecpts}
        \item Practical Example
    \end{itemize}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Nucleolus}
    \begin{itemize}
        \item Let $(N,v)$ be a $TU$ game.
        \item Let $Imp$ be the set of all imputations.
        \item The {\color{blue}$nucleolus$ $Nu(N,v)$} is the set:
        \begin{itemize}
            \item $Nu(N,v) = \{x \in Imp|\forall y \in Imp e(y)* \geq_{lex} e(x)*\}$
        \end{itemize}
    \end{itemize}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Overview}
    \begin{itemize}
     	\itemsep=.5cm
    	\item Coalition Games
    	\item Stability and Core
    	\item Fairness and Shapley Value
    	\item Other Solution Conecpts
        \item {\bf Practical Example}
    \end{itemize}
\end{frame}

@@

>>>>

\documentclass{beamer}
\usepackage[latin1]{inputenc,colortbl}
\usepackage{epsfig}
\usetheme{Frankfurt}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[page number]
%\newtheorem{definition}{Definition}
\title[Lecture 8: Coalition Games]{Lecture 8: Coalition Games}
\author{INSE 6441/2\\ \vspace{0.2cm} Applied Game Theory and Mechanism Design}
\institute{Department of Computer Science and Software Engineering\\Concordia University}
\date{Ehsan Khosrowshahi \\ \small{Acknowledgement: Stephane Airiau, Lloyd Shapley}}
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame1 title page %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\titlepage
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% frame2 outline page %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Overview}
    \begin{itemize}
     	\itemsep=.5cm
    	\item {\bf Coalition Games}
    	\item Stability and Core
    	\item Fairness and Shapley Value
    	\item Other Solution Concepts
        \item Practical Example
    \end{itemize}
\end{frame}

@@

\begin{frame}{Cooperative Game Theory}
  \begin{itemize}
     \item In game theory, a cooperative game is a game where groups of players $``coalitions''$ may enforce cooperative behaviour, hence the game is a competition between coalitions of players, rather than between individual players.
    \item Cooperation to perform a set of tasks that requires different expertise.
    \item Agents do not have enough resource on their own to perform the tasks.
    \item Examples:
    \begin{itemize}
        \item Robots have the ability to move objects in a plant, but multiple robots are required to move a heavy box.
        \item Transportation domain: agents are trucks, trains, airplanes, ships... a task is a good to be transported.
    \end{itemize}
    \item \textbf{Issues:}
        \begin{itemize}
            \item Coalition formation.
            \item Rewarding members when a task is completed.
        \end{itemize}
  \end{itemize}
\end{frame}

@@

    $\forall C \subseteq S \subseteq N$
    \begin{itemize}
        \item {\color{blue} Monotonic:} $v(C) \leq v(S)$
    \end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Solution Properties}
    Let $x \in R^n$ be a solution of the coalition game $(N,v)$
    \begin{itemize}
        \item {\color{blue} Feasible solution:} $\sum_{i \in N} x(i) \leq v(N)$
        \item {\color{blue} Efficiency:} $\sum_{i \in N} x(i) = v(N)$
        \begin{itemize}
            \item the payoff distribution is an allocation of the entire worth of the grand coalition to all agents.
        \end{itemize}
        \item {\color{blue} Individual rationality:} $\forall i \in N, x(i) \geq v(\{i\})$
        \begin{itemize}
            \item player obtains at least its self-value of payoff.
        \end{itemize}
    \end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Solution Properties (Cont.)}
    Let $x \in R^n$ be a solution of the coalition game $(N,v)$
    \begin{itemize}
        \item {\color{blue} Group rationality:} $\forall C \subseteq N, \sum_{i \in N} x(i) \geq v(C)$
        \begin{itemize}
            \item if $\sum_{i \in N} x(i) < v(C)$ is not worth it for some $C$
            \item if $\sum_{i \in N} x(i) > v(C)$ is not possible for all $C$
        \end{itemize}
%        \item {\color{blue} Pareto Optimal:} $\sum_{i \in N} x(i) = v(N)$
%        \begin{itemize}
%            \item no agent can improve its payoff without lowering the payoff of another agent
%        \end{itemize}
        \item {\color{blue} Anonymity:} a solution is independent of the names of the player.
    \end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Overview}
    \begin{itemize}
     	\itemsep=.5cm
    	\item Coalition Games
    	\item {\bf Stability and Core}
    	\item Fairness and Shapley Value
    	\item Other Solution Concepts
        \item Practical Example
    \end{itemize}
\end{frame}

@@

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} {Example 1: A two-player game}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} {Example 2: Three-player majority game}
    \begin{center}
      $N = \{1,2,3\}$ \\
      $v(\{i\}) = 0,$\\
      $v(\{C\}) = \alpha, for |C| = 2$ \\
      $v(N) = 1$ \\
    \end{center}
    \[ (x_1, x_2, x_3) \in Core(N,n) \iff \left\{
      \begin{array}{l l}
        x_i \geq 0 & \quad \forall i \in N\\
        x_i + x_j \geq \alpha & \quad \forall(i,j) \in N^2 i \neq j\\
        \sum_{i \in N} x_i = 1 &
      \end{array} \right.\]
    \[ (x_1, x_2, x_3) \in Core(N,n) \iff \left\{
      \begin{array}{l l}
        0 \leq x_i \leq 1 - \alpha & \quad \forall i \in N\\
        \sum_{i \in N} x_i = 1 &
      \end{array} \right.\]

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Overview}
    \begin{itemize}
     	\itemsep=.5cm
    	\item Coalition Games
    	\item Stability and Core
    	\item {\bf Fairness and Shapley Value}
    	\item Other Solution Concepts
        \item Practical Example
    \end{itemize}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Overview}
    \begin{itemize}
     	\itemsep=.5cm
    	\item Coalition Games
    	\item Stability and Core
    	\item Fairness and Shapley Value
    	\item {\bf Other Solution Concepts}
        \item Practical Example
    \end{itemize}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Nucleolus}
    \begin{itemize}
        \item Let $(N,v)$ be a $TU$ game.
        \item Let $Imp$ be the set of all imputations.
        \item The {\color{blue}$nucleolus$ $Nu(N,v)$} is the set:
        \begin{itemize}
            \item $Nu(N,v) = \{x \in Imp | \forall y \in Imp, e(y)^* \geq_{lex} e(x)^*\}$
        \end{itemize}
    \end{itemize}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Overview}
    \begin{itemize}
     	\itemsep=.5cm
    	\item Coalition Games
    	\item Stability and Core
    	\item Fairness and Shapley Value
    	\item Other Solution Concepts
        \item {\bf Practical Example}
    \end{itemize}
\end{frame}

@@

\begin{frame}{Experimental Results}

    \begin{figure}[!t]
    \centering
    %\includegraphics[scale=0.6]{graph1Final+.eps}
    \includegraphics[width=2.1in]{figures/task_done_opt-eps-converted-to.pdf}
    \includegraphics[width=2.1in]{figures/task_qos_opt-eps-converted-to.pdf}
    \caption{Part (a): Cumulative number of requests successfully
    answered. Part (b): Average QoS of requests performed.}
    \label{performanceall}
    \end{figure}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Improving Community Stability}
    \begin{itemize}
        \item The core may not always be non-empty, core is a strong condition.
        \item The optimal solution concepts will attract only enough services satisfying input task rate, not more.
        \item In case of web service failure, a portion of tasks will not be handled efficiently until replacement is found.
        \item By imposing cost on deviation or subsiding the community, its possible to attract more web services and keep the community stable.
        \item Will be cost efficient in case of web service failure, or for replacing web services dropping in QoS metrics as they advertised.
        \item \emph{Taxation} would need agreement between all communities (needs a centralized approach), however \emph{subsiding} can be done within communities in distributed manner.
    \end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Taxation and Subsiding}
    \begin{itemize}
        \item Least Core
        \begin{itemize}
            \item $\forall S \subseteq N, \sum_{x_i \in S} x_i \geq v(S) - \epsilon$
            \item $\epsilon$-core relaxes the core condition.
        \end{itemize}
        \item Taxation in Reliability Games \small $[Y. Bachrach, E. Elkind, 2009]$

        \item Taxation and Cost of Stability \small $[R. Meir, J. S. Rosenchein, 2011]$

        \item Taxation in Anonymous Games \small $[Y. Zick, M. Polukarov, 2013]$

        \item Relative Core
        \begin{itemize}
            \item $\lambda v(C)$ is divided among players.
            \item relative $\lambda$-Core: $\forall S \subseteq N, \sum_{x_i \in S} x_i \geq \frac{1}{\lambda}.v(S)$
            \item Clearly any community can be stabilized by large enough $\lambda$, however the community coordinator would be interested in the minimal subsidy required to stabilize the game.
            \item Can be solved using $LP$ if $v(S)$ is a linear function
        \end{itemize}
    \end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Results: Taxation and Subsiding}
    \begin{figure}[!t]
        \centering
        %\includegraphics[scale=0.6]{graph1Final+.eps}
        \includegraphics[width=3in]{figures/taxtation-eps-converted-to.pdf}
        \caption{Analysis of community subsiding coefficient $\lambda$ on
        average community size (a), cost (b), number of tasks performed
        (c), and average quality of service of tasks performed (d).}
        \label{f_taxtation}
    \end{figure}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Results: Taxation and Subsiding}
    \begin{figure}[!t]
        \centerline{\includegraphics[width=4.2in]{figures/tax_dyn-eps-converted-to.pdf}}
        \caption{Analysis of community subsiding coefficient $\lambda$
        having web service different stability levels of $\tau$ on average
        community size, number of tasks performed, average quality of
        service, and average cost/income of communities.}
        \label{fig_dynamic_taxtation}
    \end{figure}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
